\section{Summary}

\subsection{Results}

Table~\ref{tab:design-space-results} summarizes the explored design points.
The first three columns specify the architectural parameters used in each run:
input bus width, parallel input consumption, and per-layer neuron
partitioning. Several design points exceed the maximum operating frequency
typically associated with BRAM on the target device. However, these results
were achieved using LUTRAM rather than BRAM, indicating that the datapath
itself is capable of higher clock frequencies when not limited by BRAM timing.
Furthermore, all results were obtained from an out-of-context implementation,
which excludes board-level I/O and other full-system constraints.
Consequently, these frequencies should be interpreted as optimistic upper-bound
estimates rather than fully realizable system-level results. Following the
FINN-style reporting convention, FPS is computed from the steady-state
throughput rather than the end-to-end latency:
\[
\mathrm{FPS} = \frac{F_{\max}}{\text{cycles/output}}
\]
where \(F_{\max}\) is in Hz. The latency column is still reported separately as
the average cycles per image. Among the explored points, the best measured
implementation was the 256-bit / 128-input / \(64,64,10\) configuration,
which reached 15.56M FPS. The FINN paper reports a maximum of 12.3 million
image classifications per second on MNIST, so this best measured point is
1.27\(\times\) higher, or about 26.5\% above that published result. This
comparison should still be interpreted carefully, since the values reported
here come from an out-of-context LUTRAM-based implementation rather than a
full system build.

\begin{strip}
  \centering
  \captionof{table}{Current design-space results summary.}
  \label{tab:design-space-results}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{@{}rrrrrrr@{}}
      \toprule
      \textbf{Input Bus Width} & \textbf{Parallel Inputs} &
      \textbf{Parallel Neurons} & \textbf{FMAX} &
      \textbf{Cycles/Output} & \textbf{Latency} &
      \textbf{FPS} \\
      \midrule
      64   & 32  & 64,32,10      & 722 & 104.8 & 290   & 6.89M \\
      128  & 64  & 64,32,10      & 635 & 56.1  & 195.2 & 11.32M \\
      256  & 128 & 64,64,10      & 501 & 32.2  & 130.2 & 15.56M \\
      512  & 128 & 256,64,10     & --- & 17.8  & 95.2  & --- \\
      1024 & 128 & 256,64,10     & --- & 10.6  & 78.4  & --- \\
      \bottomrule
    \end{tabular}%
  }
\end{strip}

\begin{strip}
  \centering
  \captionof{table}{Resource utilization for the top three measured design points.}
  \label{tab:design-space-resources}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{@{}rrrrrrr@{}}
      \toprule
      \textbf{Input Bus Width} & \textbf{Parallel Inputs} &
      \textbf{Parallel Neurons} & \textbf{LUT} & \textbf{Registers} &
      \textbf{LUTRAM} & \textbf{BRAM} \\
      \midrule
      64   & 32  & 64,32,10  & 27723 (1.60\%) & 46757 (1.35\%)  & 8625 (0.50\%) & 0 (0.00\%) \\
      128  & 64  & 64,32,10  & 42518 (2.46\%) & 60436 (1.75\%)  & 8281 (0.48\%) & 0 (0.00\%) \\
      256  & 128 & 64,64,10  & 67663 (3.92\%) & 107002 (3.10\%) & 9037 (0.52\%) & 0 (0.00\%) \\
      \bottomrule
    \end{tabular}%
  }
\end{strip}

\paragraph{Important Judging Notes.}
Use -p 1.25 in OpenFLEX when building the design. I found this to be most consistent 
as Vivado gives up early otherwise.
Also, I was not able to get the openFLEX timing to work, as I am on a Windows device and 
the commands did not work, I tried to patch the tool to make it work on windows, but
then Modelsim would hang forever. To verify results please use make sim-original-tb.

\subsection{Implementation and Tradeoffs}

The implementation target was to maximize throughput first and then optimize
for clock frequency. Another design priority was parameterization. I wanted 
every component to be parameterized to be fully scalable to any power-of-two amount of 
parallel neurons, parallel inputs, and input width. This allowed me to gather 
these results for various configurations as I scaled up the input bus width.

Throughput is increased by replicating neuron processors
and the corresponding weight and threshold RAM banks, allowing more neurons to
be processed in parallel each cycle. However, that same replication creates
difficult timing paths because the RAM-side control and data signals become
large fanout nets with significant routing cost. To address this, multiple
pipeline stages were inserted on the interfaces that communicate with the
weight and threshold RAM units, breaking these long paths into shorter stages
and improving timing closure.

A similar approach is used in the compute reductions. The XNOR accumulation and
argmax logic are implemented as recursive balanced trees, which makes it
natural to pipeline the computation between tree levels. This reduces the logic
depth of each stage and allows the design to scale to wider buses and larger
parallel neuron counts without collapsing FMAX. The design is also heavily
parameterized, which made it possible to evaluate the architecture across a
range of bus widths, parallel-input counts, and neuron configurations while
preserving the same overall structure.

The design also supports functionality beyond what was tested for the project.
Notably, the design supports partial configuration and input buses as specified
by TKEEP, it also supports complete reconfiguration of weights and thresholds in
between image streams, as well as customizable hidden layers and topologies.

\subsection{Verification/Testbench}

The base testbench was extended into a UVM flow to verify functionality beyond
the nominal contest path and to provide more complete coverage of the design.
Some of this behavior includes: partial \texttt{TKEEP} configuration and input streams,
interleaved and reordered configuration traffic, weights-only and
thresholds-only reconfiguration, partial-layer updates, deeper hidden-layer
topologies, extreme threshold/weight/pixel values, long inter-transaction
gaps, output backpressure, and reset/reconfigure behavior.

\begin{itemize}
  \item Functional coverage was collected with\\
  \hspace*{1.5em}\texttt{make coverage-sweep}.
  \item A text coverage report was generated with\\
  \hspace*{1.5em}\texttt{make coverage-sweep-report}.
  \item The achieved-coverage report is written to\\
  \hspace*{1.5em}\texttt{coverage/}\\
  \hspace*{1.5em}\texttt{bnn\_fcc\_coverage\_sweep\_test\_coverage.txt}.
  \item Supports partial \texttt{TKEEP} configuration and input streams.
  \item Supports reconfiguring weights and thresholds after a few images.
  \item Supports interleaved configuration streams, including\\
  \hspace*{1.5em}thresholds-first and out-of-order layer updates.
  \item Supports deeper hidden-layer stacks; see\\
  \hspace*{1.5em}\texttt{make sim-ten-hidden}.
  \item Tested long delay gaps and extreme weight, threshold, and pixel values.
  \item Scales to higher neuron counts when the input bus width increases.
\end{itemize}

The coverage sweep achieved 93.18\% total covergroup coverage, covering
392 of 445 bins across 26 covergroup types, with 0 assertion failures in the
reported run.

\begin{strip}
  \centering
  \captionof{table}{Coverage summary from the directed UVM sweep.}
  \label{tab:coverage-summary}
  \small
  \renewcommand{\arraystretch}{1.18}
  \begin{tabular}{@{}p{0.18\textwidth}p{0.23\textwidth}p{0.10\textwidth}p{0.41\textwidth}@{}}
    \toprule
    \textbf{Area} & \textbf{Measured Groups} & \textbf{Status} & \textbf{Main Open Bins} \\
    \midrule
    Config stream &
    toggle 100\%, interface 100\%, handshake 100\%, packet 75.00\%, header 80.00\%, order 86.66\% &
    Strong &
    Short or single config packets, plus a few header and ordering corner bins. \\

    Image input &
    toggle 100\%, interface 100\%, handshake 100\%, spacing 100\%, image 62.50\% &
    Strong &
    Only generic short-packet image-shape bins remain; the contest image path is covered. \\

    Output stream &
    toggle 100\%, interface 100\%, backpressure 100\%, pattern 100\%, output 95.00\% &
    Strong &
    All 10 classes were observed; remaining misses are a few first-class transition bins. \\

    Config data diversity &
    weight density 100\%, threshold 86.66\% &
    Strong &
    Main miss is the threshold \texttt{unknown} ratio bin. \\

    Computational stimulus &
    output class 100\%, class transition cross 70.00\% &
    Strong &
    Every class was produced; misses are mostly ``first output was class X'' bins. \\

    Reconfiguration &
    reconfig 83.33\% &
    Good &
    Full, weights-only, thresholds-only, and partial reconfiguration were hit; the open bin is reconfiguring all layers at once. \\

    Reset / recovery &
    reset 53.75\%, post-reset 100\% &
    Partial &
    Reset timing was mostly exercised during configuration, not idle, image, output, or TLAST boundaries. \\
    \bottomrule
  \end{tabular}
  \renewcommand{\arraystretch}{1.0}
\end{strip}
