\section{Configuration Phase}

The configuration phase runs before inference and is responsible for converting
the serialized parameter stream into aligned per-layer writes. In this design,
that work is intentionally split into two levels. First, the top-level
\texttt{config\_manager} reshapes the incoming byte stream and separates weight
payloads from threshold payloads. Then each \texttt{bnn\_layer} uses a local
\texttt{config\_controller} path to distribute those writes across replicated
RAM banks.

The top-level configuration datapath is implemented in the top-level
configuration-manager RTL module. It uses four support blocks: the byte
compactor, the variable-width buffer, the message parser, and the padding FSM.
Its job is to accept the external configuration stream, compact away invalid
bytes, recover message structure, and emit write-oriented streams that are
already separated into the weight and threshold paths.

The front end of this path is byte-oriented because the external interface can
arrive with sparse valid-byte patterns. The parser then restores the higher
level meaning of the stream by decoding the message header, choosing the weight
or threshold destination, and generating the control metadata needed by the
padding logic. On the weight side, the pad state machine inserts synthetic
bytes so the downstream layer packers can emit aligned write words.

\subsection{Parser Control}

The parser in \texttt{rtl/config\_manager\_parser.sv} is responsible for
recovering message structure from the serialized byte stream. It operates as a
small three-state controller: \texttt{PARSE\_HEADER} consumes the 16-byte
header, \texttt{PARSE\_PAYLOAD} streams the payload bytes, and
\texttt{PARSE\_DONE} waits for the rest of the datapath to drain the current
message before the next header begins. In practice, this means the parser owns
all protocol-oriented responsibilities: it extracts \texttt{msg\_type} and
\texttt{layer\_id}, computes the expected payload length from the compile-time
layer tables, and asserts \texttt{payload\_start} exactly when the header has
finished.

That separation is intentional. The parser does not try to solve alignment or
packer timing at the same time it is decoding the stream. Instead, it emits a
clean payload-byte stream plus a small set of control signals and lets the next
block deal with write alignment. Splitting those responsibilities into
\texttt{config\_manager\_parser} and \texttt{config\_manager\_pad\_fsm} made the
control easier to implement and easier to reason about, because one FSM is
stream/protocol-facing while the other is write-format-facing.

\subsection{Padding Control}

The padding controller in \texttt{rtl/config\_manager\_pad\_fsm.sv} handles the
weight-side alignment problem after the parser has already identified the
message. Its state machine is likewise explicit: \texttt{IDLE} latches the
message metadata at \texttt{payload\_start}, \texttt{READ} consumes real payload
bytes, \texttt{PAD} emits synthetic \texttt{0xFF} bytes when a neuron payload
does not end on a packer-word boundary, and \texttt{DRAIN} flushes the
remaining staged bytes before the next message starts. During \texttt{PAD}, the
FSM asserts \texttt{stall} back to the parser so that the protocol side does
not advance while synthetic bytes are being inserted.

The \texttt{fifo\_vr} primitive is what makes that padded-byte path practical.
The parser first writes weight payload bytes into a byte-wide
\texttt{fifo\_weights\_bytes}, then the pad FSM chooses whether the next byte
forwarded downstream is a real FIFO byte or a synthetic padding byte. That
single-byte stream is finally repacked by the layer-local \texttt{fifo\_vr}
packer FIFOs into the wider aligned write words expected by each layer. Using
the same variable-read FIFO style for both staging and repacking keeps the byte
injection mechanism simple: the pad FSM only has to control one downstream
byte-write interface, even though the final writes leave as much wider words.

\subsection{FIFO Sizing and Backpressure}

The configuration path was also an opportunity to minimize buffering rather
than maximize throughput. Unlike the compute phase, there was no contest
performance target on the configuration stream, so the FIFOs were sized only
large enough to decouple adjacent stages and absorb the known in-flight data
already inside the pipeline. That kept the storage structures small and avoided
adding unnecessary area or deeper buffer-management logic to a path that only
needs to configure the design before inference.

The main place where this matters is the byte-serialization FIFO inside
\texttt{rtl/config\_manager.sv}. The external stream first arrives in wide AXI
beats, then gets compacted and repacked before the parser consumes only one
byte per cycle. That width conversion creates the real burst-mismatch point in
the design, so this serializer FIFO is the only configuration buffer that
needs meaningful headroom. Even there, the design still keeps the FIFO small
and relies on early backpressure: \texttt{config\_ready} is driven from the
FIFO's almost-full flag rather than from its true full flag. The almost-full
threshold is set with an explicit reserve for the three whole write words that
can still be in flight after \texttt{config\_ready} deasserts: one word
buffered in \texttt{vw\_buffer}, one in the local write pipeline, and one more
still emerging from the registered byte compactor. By stalling one write slot
early, the design guarantees those words still fit without having to over-size
the FIFO.

The downstream buffers are kept even smaller because they are only local
decoupling points. The weight-byte FIFO is just a short queue between the
parser and the pad FSM, the threshold FIFO only needs enough storage to repack
threshold bytes into words, and each layer packer FIFO holds only a very small
number of aligned output words. In other words, the design uses just enough
elasticity to tolerate local staging differences, then leans on almost-full
driven upstream throttling and explicit FSM stall signals to keep the
configuration stream safe instead of paying for large buffers everywhere.

\clearpage
\begin{strip}
  \centering
  \resizebox{!}{0.73\textheight}{\input{figures/config_manager_diagram}}
  \captionof{figure}{Block diagram of the \texttt{config\_manager} datapath and
  its primary interface signals.}
  \Description{Vertical block diagram of the configuration manager showing
  byte compaction, repacking, parsing, weight padding, threshold packing, and
  the external write interfaces that leave the module.}
  \label{fig:config-manager}
\end{strip}

\subsection{Layer-Local Write Distribution}

After the top-level parser finishes that global reshaping step, each
\texttt{bnn\_layer} still has to fan the write stream out to replicated local
memories. That is the purpose of \texttt{config\_controller} and the
surrounding staging registers in \texttt{rtl/bnn\_layer.sv}. The controller
generates the bank-select and address signals, while the aligned data stream
bypasses the controller datapath and is retimed for fanout.

Those fanout pipeline registers are a central part of the implementation, not a
minor detail. They cut the long controller-to-BRAM distribution path into a
manageable stage while preserving a steady stream of writes after the pipeline
fills. That is why this report treats the local write-distribution block as its
own step in the overall configuration phase.

\clearpage
\begin{strip}
  \centering
  \resizebox{0.90\textwidth}{!}{\input{figures/config_controller_diagram}}
  \captionof{figure}{Layer-local config-write distribution around
  \texttt{config\_controller}. The light gray fanout pipeline registers mark
  the timing-cut stage that drives the per-bank weight and threshold RAMs.}
  \Description{Block diagram of the layer-local configuration write path
  showing staged weight and threshold inputs, the config controller, timing-cut
  fanout registers, and replicated RAM banks for weight and threshold storage.}
  \label{fig:config-controller-block}
\end{strip}

\subsection{Timing Optimizations}

This layer-local write-distribution block turned into one of the important
timing problems in the overall design because it sits between a relatively
compact control source and a wide set of physically distributed RAM banks. As
the neuron count grows, the weight and threshold write enables, bank selects,
addresses, and data all have to fan out across many destinations. That makes
the path sensitive not only to logic depth, but also to placement and routing
distance.

The solution used here was to instantiate the staging registers explicitly and
preserve them with \texttt{keep}. Without that constraint, synthesis tended to
absorb or merge the hand-placed pipeline stages back into neighboring logic,
which removed the intended timing cut and recreated the same long fanout path.
By keeping those registers structurally visible, the implementation preserves a
clear controller stage followed by a separate fanout stage.

In practice, two register stages are used. The first stage captures the
incoming write stream and controller-facing control signals near the local
configuration ingress. The second stage sits at the RAM-distribution boundary
and re-drives the per-bank write signals toward the replicated memories. That
extra stage matters because the RAM banks can end up placed far apart after
implementation; they are not guaranteed to remain close to the controller or to
each other. With the additional retiming point, the placer and router have much
more freedom to cluster the final fanout stage near the RAM banks, which is why
this path no longer dominates the critical timing after optimization.

\clearpage
