\section{Compute Phase}

Once the model parameters are loaded, the design switches to the compute phase.
This phase again has two main pieces. The first is the input-preparation path
in \texttt{data\_in\_manager}, which converts the incoming activation stream
into the binary words expected by the network. The second is the layer-local
forward datapath inside \texttt{bnn\_layer}, where buffered activations, RAM
reads, and neuron-processor outputs are coordinated.

\subsection{Input Framing and Packing}

\texttt{rtl/data\_in\_manager.sv} mirrors several of the same structural ideas
used in the configuration path, but now for activation traffic rather than
parameter traffic. The front end again begins with
\texttt{tkeep\_byte\_compactor}, because the image stream can arrive with sparse
valid-byte patterns. After compaction, local frame-count logic tracks how many
real bytes have been accepted in the current image and determines whether the
final beat needs zero padding so the frame length becomes a clean multiple of
\texttt{PARALLEL\_INPUTS}. That merged real-data/padding stream is then written
through \texttt{vw\_buffer}, which repacks the byte stream into full internal
words.

This block is intentionally separate from the actual layer compute logic. By
finishing byte compaction, frame shaping, and word packing at the top level,
the \texttt{bnn} core sees a regular stream instead of having to carry AXI byte
validity and end-of-frame corner cases deep into the runtime datapath.

\subsection{Binarization and Staging}

After repacking, \texttt{data\_in\_manager} converts each byte into one binary
activation bit using the local binarization threshold and writes those bits
into the internal \texttt{bin\_fifo}. This FIFO is more than a convenience
buffer: it decouples the image-side repacking path from the BNN-side
\texttt{bnn\_ready}/\texttt{bnn\_en} handshake, and it provides the backpressure
signal that throttles the upstream byte path before the staging queue
overflows. In other words, the compute phase still starts with a small stream
conditioning problem, and the FIFO is what isolates that problem from the
first-layer datapath.

\clearpage
\begin{strip}
  \centering
  \resizebox{!}{0.72\textheight}{\input{figures/data_in_manager_diagram}}
  \captionof{figure}{Block diagram of the \texttt{data\_in\_manager} datapath
  and its primary interface signals. The light gray staging block marks the
  BNN-facing queue that decouples input reshaping from the core compute flow.}
  \Description{Block diagram of the data input manager showing valid-byte
  compaction, frame shaping, repacking, binarization, and the staged interface
  that feeds the BNN core.}
  \label{fig:data-in-manager}
\end{strip}
\clearpage

\clearpage
\begin{strip}
  \begin{minipage}[t]{0.48\textwidth}
    \subsection{Replay and Input Alignment}

    Inside \texttt{rtl/bnn\_layer.sv}, the forward path begins with the local
    \texttt{replay\_buffer}. This block allows one accepted activation frame to
    be reused across the repeated neuron-group computations required by the
    layer, instead of demanding the external source resend the same activations
    for every parallel-neuron slice. The replayed activation word then passes
    through the explicit alignment delay stage so it reaches the neuron
    processors in step with the parameter-read pipeline.

    That alignment stage matters because the RAM path is not a single-cycle
    lookup. Between the controller scheduling, local per-bank address registers,
    and the registered RAM outputs, the weight and threshold operands emerge
    several cycles after the replay buffer launches the activation word.
    Keeping the input side explicitly delayed in the diagram makes that timing
    relationship much easier to understand than burying it inside the
    controller or RAM descriptions.
  \end{minipage}\hfill
  \begin{minipage}[t]{0.48\textwidth}
    \subsection{Read Control and Neuron Array}

    The second half of the compute phase is the parameter-read and evaluation
    path. \texttt{neuron\_controller} sequences the shared weight and threshold
    read addresses, while the banked RAM structures provide one local operand
    stream per parallel neuron. The gray read-register stages shown in the
    diagram represent the staging/alignment points that keep the controller
    side, RAM side, and neuron-processor side synchronized. Finally, the
    \texttt{u\_np[gi]} array consumes aligned activations, weights, and
    thresholds and produces the next layer's output bits along with the
    accumulator counts used by the last-stage argmax tree.

    As in the configuration phase, the implementation is easier to follow
    because the responsibilities are divided cleanly: replay and alignment
    handle the activation side, the controller and RAM banks handle operand
    scheduling, and the neuron processors perform the actual
    XNOR-accumulate-and-threshold work.
  \end{minipage}

  \vspace{0.75em}

  \centering
  \resizebox{0.86\textwidth}{!}{\input{figures/bnn_layer_compute_diagram}}
  \captionof{figure}{Simplified compute flow inside one \texttt{bnn\_layer}.
  The light gray blocks mark the explicit staging and alignment registers in the
  activation and read-data paths.}
  \Description{Block diagram of the layer compute flow showing replay buffering
  of activations, alignment, controller-driven reads from weight and threshold
  RAM banks, and the final neuron-processor array that produces outputs.}
  \label{fig:bnn-layer-compute}
\end{strip}

\subsection{Recursive Reduction Trees}

Inside the neuron datapath, the XNOR-popcount reduction is implemented with the
recursive pipelined tree in \texttt{rtl/xnor\_add\_tree.sv}, which then hands
its partial counts to the recursive \texttt{rtl/add\_tree.sv} reduction stage.
The structure was inspired by Greg Stitt's recursive adder-tree approach on
StittHub, but adapted here for the binary XNOR-accumulate workload used by the
BNN layer. The main benefit of that recursive structure is that it naturally
builds a balanced pipelined reduction instead of a long linear accumulation
chain, so the reduction depth grows logarithmically with the parallel-input
count.

There is also one more explicit register cut after the tree output inside
\texttt{rtl/neuron\_processor.sv}. When the last partial sum for a neuron
arrives, the final accumulated count and aligned threshold are first captured
locally, and only in the following cycle does the design perform the
threshold-compare operation. That extra register stage splits the reduction path
from the comparison logic so the tree output plus compare no longer form one
combined critical path.

The output block uses the same general idea. The class-score reduction in
\texttt{rtl/argmax\_tree.sv} is also implemented as a recursive pipelined tree,
with generated alignment delays on the shorter branch and a registered compare
at each merge point.
\clearpage
