\section{Verification}

The UVM environment was built by extending the existing directed
\texttt{bnn\_fcc\_tb} flow rather than replacing it. The original testbench
already had the important design-specific assets: the 
\texttt{BNN\_FCC\_Model}, configuration encoding, image stimulus generation,
probability-based valid throttling, randomized output backpressure, and
end-to-end pass/fail checking. The UVM work reorganized those capabilities into
reusable layers so new tests could vary configuration order, partial
reconfiguration, \texttt{TKEEP} packetization, reset timing, and image
selection without rewriting low-level AXI signal code.

\subsection{Top-Level UVM Wrapper}

The top wrapper in \texttt{bnn\_fcc\_uvm\_tb.sv}
preserves the same runtime controls as the original testbench: topology
selection, image count, model-verification enable, valid probabilities,
backpressure mode, timeout, and clock period. It instantiates the DUT together
with the three AXI4-Stream interfaces for configuration input, image input, and
output, then publishes those interfaces through \texttt{uvm\_config\_db}.

One useful addition in the UVM flow is \texttt{bnn\_fcc\_ctrl\_if}. That shared
control interface gives tests, the scoreboard, and the coverage components a
clean place to coordinate reset and output-ready forcing without directly
reaching into the top module internals. The wrapper also keeps the output-side
protocol assertions local to the signal layer, while the scoreboard handles the
end-to-end functional checks.

\subsection{Environment and Agents}

The main environment in \texttt{bnn\_fcc\_env.sv}
assembles three protocol agents, one scoreboard, and four coverage collectors.
The configuration and image agents are active AXI4-Stream masters, while the
output agent is passive because the DUT owns the output channel. This mirrors
the real interface roles and keeps the output monitoring separate from output
ready generation.

The environment forces all three monitors into packet-level mode before
connecting them. That decision matters because the scoreboard and coverage
components reason about full configuration messages, full image packets, and
full outputs rather than isolated beats. The environment then connects each
monitor to both the scoreboard and the configuration, input, output, and
cross-stream system coverage blocks.

\subsection{Sequence Architecture}

Most of the reuse in the UVM flow lives in the shared sequence library
\texttt{bnn\_fcc\_sequences.svh}. The configuration side is built around one
common base sequence that reuses the shared \texttt{BNN\_FCC\_Model} to
generate exactly the same legal configuration messages as the original
testbench. Its key knobs select whether weights and thresholds are included,
which layers are touched, and what message ordering is used. Together those
controls cover the important reconfiguration scenarios: full configuration,
partial configuration, weights-only updates, thresholds-only updates, and
reordered traffic such as thresholds-first or layer-interleaved streams.

The image side follows the same pattern. A shared base image sequence loads
stimulus from the same reference path used by the directed testbench, or
generates randomized images for custom topologies, then packs those images into
AXI beats. Thin derived classes change only the presentation style: beat-level
traffic, packet-level traffic, scripted image-index traffic, scripted raw-pixel
traffic, and \texttt{TKEEP}-repacked packet traffic. This is what keeps
derived tests short: they choose a scenario, but they do not reimplement model
knowledge or packet formatting.

\subsection{Scoreboard and Reset Handling}

The scoreboard in \texttt{bnn\_fcc\_scoreboard.svh} is
state-aware because this design supports both reconfiguration and reset during
test execution. Instead of comparing outputs immediately at the moment they
appear, the scoreboard queues expected predictions when input packets arrive.
Each queued entry stores the expected class, the logical image index, and a
configuration epoch. Tests call \texttt{commit\_model()} after sending a new
configuration, and the scoreboard copies that model into an internal snapshot.
That protects the checks from later test-side edits to the shared model handle.

Reset handling is also explicit. When reset asserts through
\texttt{ctrl\_if.rst}, the scoreboard clears all pending expectations, marks
itself unconfigured, and ignores later outputs until a fresh post-reset model
is committed. This avoids stale pre-reset expectations being compared against
post-reset outputs. The same scoreboard also tracks average latency and average
throughput using handshake timestamps reconstructed by the packet monitors.

\subsection{Coverage-Oriented Tests}

Because the infrastructure is layered, focused tests stay short. A test only
needs to choose a configuration sequence flavor, an image sequence flavor,
timing stress knobs, and optional reset or reconfiguration events. The shared
environment handles the agent connections, the scoreboard model snapshots, and
the coverage collection. The dedicated \texttt{bnn\_fcc\_coverage\_sweep\_test}
then serializes many of these directed scenarios into one judge-friendly run,
producing one simulation log, one UCDB, and one text coverage report.

The result is a verification flow that is specific to \texttt{bnn\_fcc} rather
than a generic UVM demo. The original model and stimulus logic are still the
source of truth, but UVM turns them into reusable protocol-layer components,
scenario-driven sequences, and a reset-aware scoreboard that can support a much
broader set of verification scenarios than the original monolithic testbench.
